---
title: Scraping data from WHO-DON website
author: Luong Nguyen Thanh
---

## Load necessary packages

```{r}
pacman::p_load(
  rio,
  here,
  httr,
  jsonlite,
  stringr,
  tidyverse)
```

## Write a function to get data from WHO-DON website

Using method GET from `httr` package to get data from WHO-DON website. The function `get_who_news` will take a URL as input and return the data from the API. The function will return `NULL` if the request was unsuccessful.

Status code = 200 means the connection is successful. The function will parse the JSON content and return the data.

Next, we will initialize the variables and loop to fetch all pages. The function will check if there is a next page link and keep fetching until there is no next page link.

```{r}
#| message: FALSE
#| warning: FALSE
#| eval: FALSE

# Function to get news data from a specific URL
get_who_news <- function(url) {
  # Make the GET request to the API
  response <- GET(url)
  
  # Check the status of the response
  if (status_code(response) == 200) {
    # Parse the JSON content
    content <- content(response, "text")
    data <- fromJSON(content)
    return(data)
  } else {
    # Return NULL if the request was unsuccessful
    return(NULL)
  }
}

# Initialize variables
base_url <- "https://www.who.int/api/news/diseaseoutbreaknews"
all_news <- list()
next_url <- base_url
keep_fetching <- TRUE

# Loop to fetch all pages
while (keep_fetching) {
  data <- get_who_news(next_url)
  
  if (!is.null(data) && "value" %in% names(data)) {
    all_news <- c(all_news, data$value)
    
    # Check if there is a next page link
    if ("@odata.nextLink" %in% names(data)) {
      next_url <- data[["@odata.nextLink"]]
    } else {
      keep_fetching <- FALSE
    }
  } else {
    keep_fetching <- FALSE
  }
}
```

## Convert data from list to wide dataframe

The data is currently stored as a nested list. We will convert this nested list to a wide data frame for further analysis. We will define a function `convert_to_df` that takes the nested list as input and returns a data frame.

Some cleaning steps are performed to remove HTML tags from the text data.

Finally, we will export the data frame to a CSV file for further analysis.

```{r}
#| eval: FALSE
# Define a function to convert the nested list to a data frame
convert_to_df <- function(news_list) {
  # Initialize an empty list to hold data frames
  df_list <- list()
  
  # Determine the segment length
  segment_length <- 22
  
  # Iterate through the list in steps of segment_length
  for (i in seq(1, length(news_list), by = segment_length)) {
    # Extract the current segment
    segment <- news_list[i:(i + segment_length - 1)]
    
    # Convert the segment to a data frame and add it to the list
    df_list[[length(df_list) + 1]] <- as.data.frame(segment)
  }
  
  # Combine all data frames into one
  combined_df <- do.call(rbind, df_list)
  return(combined_df)
}

# Function to remove HTML tags
remove_html_tags <- function(text) {
  return(str_replace_all(text, "<[^>]*>", ""))
}

all_news_df <- convert_to_df(all_news)


all_news_df %>% 
  mutate(across(where(is.character), remove_html_tags)) %>%
  arrange(desc(PublicationDate)) %>%
  rio::export(here("data", "who_dons.csv"))
```


## Pivot data into long format

```{r}
corpus <- import(here("data/who_dons.csv")) %>% 
    select(
        UrlName,
        DonId,
        # Ranking of information based on important
        Summary,
        Overview,
        Epidemiology,
        Assessment,
        Advice,
        FurtherInformation
    ) %>% 
    # remove value 1 in character columns to NA
    mutate(across(where(is.character), ~na_if(., ""))) %>% 
    # Standardize DonID column, with this format Year-DON-Number
    # the UrlName column will be used if DonID is missing
    mutate(DonID_standardized = coalesce(DonId, UrlName)) %>%
    #relocate DonID_standardized column to the first column
    relocate(DonID_standardized, .before = UrlName) %>%
    # pivot_longer data, so information from columns summary to FurtherInformation are in one column
    pivot_longer(
        cols = Summary:FurtherInformation,
        names_to = "InformationType",
        values_to = "Text") 

export(corpus, here("data", "corpus.csv"))
```

## Preview data

```{r}
#| echo: FALSE
all_news_df <- import(here("data", "who_dons.csv"))
```

Data up to 15 July 2024

```{r}
#| label: glimpse-data
glimpse(corpus)
```
